%===================================== CHAP 4 =================================

\chapter{Experiments and results}\label{chpt:experiments}

\section{Plan}

\section{Setup}

hpc = HPC([io\_dim, 240, 1600, 480, io\_dim], \\
          0.67, 0.25, 0.04,  \# connection rates: (in\_ec, ec\_dg, dg\_ca3) \\
          0.10, 0.01, 0.04,  \# firing rates: (ec, dg, ca3) \\
          0.7, 100.0, 0.1, turnover\_rate,  \# gamma, epsilon, nu, turnover rate \\
          0.10, 0.95, 0.8, 2.0, weighting\_dg)  \# k\_m, k\_r, a\_i, alpha. alpha is 2 in 4.1
          
\section{Experiments}
\subsection{Experiment pre: Model calibration}
\subsection{... Results}

\subsection{Experiment 1: Consolidation performance}
\subsection{Results experiment 1: consolidation performance}

\subsection{Experiment Y: Novel}



% ========================== section ============================
Experiment design
Results
Comparisons

\section{Preliminary results}

\textbf{100 \% connection ratio EC-CA3:}

Fairly rapid convergence for three patterns in HPC-module for turnover between every training set iteration. 
Not necessarily successful recall of all patterns. Does this have something to do with the synchronized CA3-layer during recall? Separation possible during recall when the desired pattern(s) are presented to the network - however, not all may be recalled.

-> New random pattern each time stability was reached resulted in better recall.

Is this also the case for heavier weighting of the DG-CA3 path during learning?

Spurious pattern reduction/correlation with occurrence when using turnover?

Convergence when turnover is removed between set iterations?

\section{Model calibration}

Experiments designed for model calibration

Dimensions analyzed outlined above.

First: STM-network extraction rate (at first, empirically observed to be same as solely auto-associative Hopfield network).


\textbf{Notes}

experiments suite - two as outlined by \citep{Hattori2014}, originally retrieved from ... as outlined above
enabling several trials automatically.

Turnover between every training set iteration (?). Needs to include empirical data on decision making. Move to preliminary experimentation in chpt. 4?

\cleardoublepage